---
title: "Untitled"
output: html_document
---
https://csgillespie.github.io/efficientR/introduction.html

# 1. Introduction
1.6.2 Benchmarking example
```{r}
library("microbenchmark")
df = data.frame(v = 1:4, name = letters[1:4])
microbenchmark(df[3, 2], df[3, "name"], df$name[3])
# Unit: microseconds
#          expr     min    lq  mean median    uq   max neval cld
#      df[3, 2]   17.99 18.96 20.16  19.38 19.77 35.14   100   b
# df[3, "name"]   17.97 19.13 21.45  19.64 20.15 74.00   100   b
#    df$name[3]   12.48 13.81 15.81  14.48 15.14 67.24   100   a # BEST
```
1.6.3 Profiling

```{r}
#install.packages("profvis")
library("profvis")
profvis(expr = {

  # Stage 1: load packages
  # library("rnoaa") # not necessary as data pre-saved
  library("ggplot2")

  # Stage 2: load and process data
  out = readRDS("extdata/out-ice.Rds")
  df = dplyr::rbind_all(out, id = "Year")

  # Stage 3: visualise output
  ggplot(df, aes(long, lat, group = paste(group, Year))) +
    geom_path(aes(colour = Year))
  ggsave("figures/icesheet-test.png")
}, interval = 0.01, prof_output = "ice-prof")
```
# 2 Efficient set-up
```{r}
#install.packages("benchmarkme")
library("benchmarkme")
```


```{r}
Sys.info()
```
# 3 Efficient programming

```{r}
library("compiler")
library("memoise")
```

Top 5 tips for efficient programming

- Be careful never to grow vectors.
- Vectorise code whenever possible.
- Use factors when appropriate.
- Avoid unnecessary computation by caching variables.
- Byte compile packages for an easy performance boost.

```{r}
method1 = function(n) {
  vec = NULL # Or vec = c()
  for (i in seq_len(n))
    vec = c(vec, i)
  vec
} # slow

method2 = function(n) {
  vec = numeric(n)
  for (i in seq_len(n))
    vec[i] = i
  vec
} # middle

method3 = function(n) seq_len(n) # very fast
```


```{r}
n=10000
microbenchmark(times = 100, unit = "s",
               method1(n), method2(n), method3(n))
```


```{r}
method3(100)
```


```{r}
x = runif(n) + 1
x
```

3.2.2 Vectorised code
```{r}
n=1000000
x = runif(n) + 1
```

```{r}
#install.packages("tictoc")
library(tictoc)
```

Compare for loop and sum:
```{r}
# method 1
tic()
log_sum = 0
for (i in 1:length(x))
  log_sum = log_sum + log(x[i])
toc() # 0.17
```


```{r}
# method 2 (VERY fast)
tic()
log_sum = sum(log(x))
toc() #0.06 (faster than 3 times)
```
Compare montecarlo for loop and sum

```{r}
monte_carlo = function(N) {
  hits = 0
  for (i in seq_len(N)) {
    u1 = runif(1)
    u2 = runif(1)
    if (u1 ^ 2 > u2)
      hits = hits + 1
  }
  return(hits / N)
}
```


```{r}
N = 500000
system.time(monte_carlo(N)) # 2.58 s
```

```{r}
tic()
monte_carlo_vec = function(N) sum(runif(N)^2 > runif(N)) / N
toc() # 0.02 s
```
```{r}
install.packages("devtools")
devtools::install_github("csgillespie/efficientR")
```


```{r}

data("ex_mat", package = "efficient")
```

# https://www.r-bloggers.com/2016/01/strategies-to-speedup-r-code/

```{r}
# Create the data frame
col1 <- runif (12^5, 0, 2)
col2 <- rnorm (12^5, 0, 2)
col3 <- rpois (12^5, 3)
col4 <- rchisq (12^5, 2)
df <- data.frame (col1, col2, col3, col4)
df
```


```{r}
# Original R code: Before vectorization and pre-allocation
system.time({
  for (i in 1:nrow(df)) { # for every row
    if ((df[i, 'col1'] + df[i, 'col2'] + df[i, 'col3'] + df[i, 'col4']) > 4) { # check if > 4
      df[i, 5] <- "greater_than_4" # assign 5th column
    } else {
      df[i, 5] <- "lesser_than_4" # assign 5th column
    }
  }
})
```
Vectorise and pre-allocate data structures
```{r}
# after vectorization and pre-allocation
output <- character (nrow(df)) # initialize output vector
system.time({
  for (i in 1:nrow(df)) {
    if ((df[i, 'col1'] + df[i, 'col2'] + df[i, 'col3'] + df[i, 'col4']) > 4) {
      output[i] <- "greater_than_4"
    } else {
      output[i] <- "lesser_than_4"
    }
  }
df$output})
```


```{r}
#install.packages("ggplot2movies")
library("profvis")
library(ggplot2movies)
profvis({
  data(movies, package = "ggplot2movies") # Load data
  movies = movies[movies$Comedy == 1,]
  plot(movies$year, movies$rating)
  model = loess(rating ~ year, data = movies) # loess regression line
  j = order(movies$year)
  lines(movies$year[j], model$fitted[j]) # Add line to the plot
})
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

```{r}
```

